\begin{thebibliography}{33}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\providecommand{\urlprefix}{}

\bibitem[{Agrawal et~al.(1993)Agrawal, Imieli\'{n}ski, and Swami}]{AgrawalIS93}
Agrawal, R., Imieli\'{n}ski, T., Swami, A.: Mining association rules between
  sets of items in large databases.
\newblock SIGMOD Rec. 22, 207--216 (1993)%,
  %\urlprefix\url{http://doi.acm.org/10.1145/170036.170072}

\bibitem[{Alon and Spencer(2008)}]{AlonS08}
Alon, N., Spencer, J.H.: The Probabilistic Method.
%\newblock Interscience Series in Discrete Mathematics and Optimization, John
%  Wiley {\&} Sons, Hoboken, NJ, USA, third edn. (2008)
\newblock Wiley, third edn. (2008)

\iffalse
\bibitem[{Bolton et~al.(2002)Bolton, Hand, and Adams}]{BoltonHA02}
Bolton, R.J., Hand, D.J., Adams, N.M.: Determining hit rate in pattern search.
%\newblock In: Pattern Detection and Discovery. Lecture Notes in Computer
%  Science, vol. 2447, pp. 36--48 (2002)
\newblock In: Pattern Detection and Discovery. LNCS, vol. 2447, pp. 36--48 (2002)
\fi

\bibitem[{Boucheron et~al.(2005)Boucheron, Bousquet, and
  Lugosi}]{BoucheronBL05}
Boucheron, S., Bousquet, O., Lugosi, G.: Theory of classification : A survey of
  some recent advances.
\newblock {ESAIM}: Probab.~and Statistics 9, 323--375 (2005)%,
  %\urlprefix\url{http://cat.inist.fr/?aModele=afficheN\&\#38;cpsidt=17367966}

  \iffalse
\bibitem[{Brijs et~al.(1999)Brijs, Swinnen, Vanhoof, and Wets}]{BrijsSVW99}
Brijs, T., Swinnen, G., Vanhoof, K., Wets, G.: Using association rules for
  product assortment decisions: A case study.
%\newblock In: Fayyad, U.M., Chaudhuri, S., Madigan, D. (eds.) KDD. pp.
%  254--260. ACM (1999)
\newblock KDD'99 (1999)
\fi

\iffalse
\bibitem[{Dudoit et~al.(2003)Dudoit, Shaffer, and Boldrick}]{DudoitSB03}
Dudoit, S., Shaffer, J., Boldrick, J.: Multiple hypothesis testing in
  microarray experiments.
\newblock Statistical Science 18(1), 71--103 (2003)
\fi

\bibitem[{DuMouchel and Pregibon(2001)}]{DuMouchelP01}
DuMouchel, W., Pregibon, D.: Empirical {B}ayes screening for multi-item
  associations.
%\newblock In: Proceedings of the seventh ACM SIGKDD international conference on
%  Knowledge discovery and data mining. pp. 67--76. KDD '01, ACM, New York, NY,
%  USA (2001), %\urlprefix\url{http://doi.acm.org/10.1145/502512.502526}
\newblock KDD'01 (2001)

\bibitem[{Geng and Hamilton(2006)}]{GengH06}
Geng, L., Hamilton, H.J.: Interestingness measures for data mining: A survey.
%\newblock ACM Computing Surveys 38(3) (2006),
\newblock ACM Comp.~Surv.~38(3) (2006)%,
  %\urlprefix\url{http://doi.acm.org/10.1145/1132960.1132963}

  \iffalse
\bibitem[{Geurts et~al.(2003)Geurts, Wets, Brijs, and Vanhoof}]{GeurtsWBV03}
Geurts, K., Wets, G., Brijs, T., Vanhoof, K.: Profiling of high-frequency
  accident locations by use of association rules.
\newblock Transp.~Res.~Rec.: J.~of the Transp.~Res.~Board 1840, 123--130 (2003)
\fi

\bibitem[{Gionis et~al.(2007)Gionis, Mannila, Mielik\"{a}inen, and
  Tsaparas}]{GionisMMT07}
Gionis, A., Mannila, H., Mielik\"{a}inen, T., Tsaparas, P.: Assessing data
  mining results via swap randomization.
%\newblock ACM Transactions on Knowledge Discovery from Data 1(3) (Dec 2007),
\newblock ACM Trans.~on Knowl.~Disc.~from Data 1(3) (2007)%,
  %\urlprefix\url{http://doi.acm.org/10.1145/1297332.1297338}

\bibitem[{Goethals and Zaki(2004)}]{GoethalsZ04}
Goethals, B., Zaki, M.J.: Advances in frequent itemset mining implementations:
  report on {FIMI'03}.
  %\newblock SIGKDD Explor. Newsl. 6(1), 109--117 (Jun 2004),
\newblock SIGKDD Explor. Newsl. 6(1), 109--117 (2004)%,
  %\urlprefix\url{http://doi.acm.org/10.1145/1007730.1007744}

\bibitem[{Goldshmidt et~al.(1994)Goldshmidt, Nehme, and Yu}]{GoldschmidtNY94}
Goldshmidt, O., Nehme, D., Yu, G.: Note: On the set-union knapsack problem.
\newblock Naval Research Logistics 41(6), 833--842 (1994)

\bibitem[{Grahne and Zhu(2003)}]{GrahneZ03}
Grahne, G., Zhu, J.: Efficiently using prefix-trees in mining frequent
  itemsets.
%\newblock In: Goethals, B., Zaki, M.J. (eds.) FIMI. CEUR Workshop Proceedings,
%  vol.~90. CEUR-WS.org (2003)
\newblock FIMI'03 (2003)

\bibitem[{H{\"a}m{\"a}l{\"a}inen(2010)}]{Hamalainen10}
H{\"a}m{\"a}l{\"a}inen, W.: {StatApriori}: an efficient algorithm for searching
  statistically significant association rules.
%\newblock Knowledge and Information Systems 23(3), 373--399 (2010)
\newblock Knowl.~and Inf.~Sys.~23(3), 373--399 (2010)

\bibitem[{Han et~al.(2007)Han, Cheng, Xin, and Yan}]{HanCXY07}
Han, J., Cheng, H., Xin, D., Yan, X.: Frequent pattern mining: current status
  and future directions.
\newblock Data Min.~and Knowl.~Disc.~15, 55--86 (2007),
  %\urlprefix\url{http://dx.doi.org/10.1007/s10618-006-0059-1}

\bibitem[{Han et~al.(2006)Han, Kamber, and Pei}]{han2006data}
Han, J., Kamber, M., Pei, J.: Data mining: concepts and techniques.
\newblock Morgan Kaufmann (2006)

\bibitem[{Hanhij{\"a}rvi(2011)}]{Hanhijarvi11}
Hanhij{\"a}rvi, S.: Multiple hypothesis testing in pattern discovery.
%\newblock In: Discovery Science. Lecture Notes in Computer Science, vol. 6926,
%  pp. 122--134 (2011)
\newblock DS'11 (2011)

\bibitem[{Har-Peled and Sharir(2011)}]{HarPS11}
Har-Peled, S., Sharir, M.: Relative $(p,\varepsilon)$-approximations in
  geometry.
\newblock Discr.~\& Comput.~Geom.~45(3), 462--496 (2011)
  %\urlprefix\url{http://dx.doi.org/10.1007/s00454-010-9248-1}

\iffalse
\bibitem[{Jacquemont et~al.(2009)Jacquemont, Jacquenet, and
  Sebban}]{JacquemontFS09}
Jacquemont, S., Jacquenet, F., Sebban, M.: A lower bound on the sample size
  needed to perform a significant frequent pattern mining task.
\newblock Patt.~Recog.~Lett.~30(11), 960--967 (2009),
  %\urlprefix\url{http://www.sciencedirect.com/science/article/pii/S0167865509000968}
\fi

\iffalse
\bibitem[{Jaroszewicz et~al.(2009)Jaroszewicz, Scheffer, and
  Simovici}]{JaroszewiczSS09}
Jaroszewicz, S., Scheffer, T., Simovici, D.A.: Scalable pattern mining with
  bayesian networks as background knowledge.
%\newblock Data Mining and Knowledge Discovery 18(1), 56--100 (2009)
\newblock Data Min.~and Knowl.~Disc.~18(1), 56--100 (2009)
\fi

\bibitem[{Kirsch et~al.(2012)Kirsch, Mitzenmacher, Pietracaprina, Pucci, Upfal,
  and Vandin}]{KirschMAPUV12}
Kirsch, A., Mitzenmacher, M., Pietracaprina, A., Pucci, G., Upfal, E., Vandin,
  F.: An efficient rigorous approach for identifying statistically significant
  frequent itemsets.
\newblock J.~ACM 59(3), 12:1--12:22 ( 2012)
  %\urlprefix\url{http://doi.acm.org/10.1145/2220357.2220359}

\iffalse
\bibitem[{Lallich et~al.(2006)Lallich, Teytaud, and Prudhomme}]{LallichTP06}
Lallich, S., Teytaud, O., Prudhomme, E.: Statistical inference and data mining:
  false discoveries control.
%\newblock In: Compstat 2006 - Proceedings in Computational Statistics. pp.
%  325--336 (2006)
\newblock Compstat'06 (2006)
\fi

\iffalse
\bibitem[{Li et~al.(2001)Li, Long, and Srinivasan}]{LiLS01}
Li, Y., Long, P.M., Srinivasan, A.: Improved bounds on the sample complexity of
  learning.
\newblock J.~of Comp.~and Sys.~Sci.~62(3), 516--527 (2001)
  %\urlprefix\url{http://www.sciencedirect.com/science/article/B6WJ0-45B65MX-H/2/3b0f2dd50a30dcee8b995d701959f15f}
\fi

\bibitem[{Liu et~al.(2011)Liu, Zhang, and Wong}]{LiuZW11}
Liu, G., Zhang, H., Wong, L.: Controlling false positives in association rule
  mining.
\newblock VLDB'11 (2011)
%\newblock Proc.~of the VLDB Endow.~5(2), 145--156 (2011),
  %\urlprefix\url{http://dl.acm.org/citation.cfm?id=2078324.2078330}

\bibitem[{L\"{o}ffler and Phillips(2009)}]{LofflerP09}
L\"{o}ffler, M., Phillips, J.M.: Shape fitting on point sets with probability
  distributions.
\newblock ESA'09 (2009)
%\newblock In: Fiat, A., Sanders, P. (eds.) Algorithms - ESA 2009, Lecture Notes
%  in Computer Science, vol. 5757, pp. 313--324. Springer, Berlin / Heidelberg
%  (2009), %\urlprefix\url{http://dx.doi.org/10.1007/978-3-642-04128-0_29}

\bibitem[{Megiddo and Srikant(1998)}]{MegiddoS98}
Megiddo, N., Srikant, R.: Discovering predictive association rules.
%\newblock In: Agrawal, R., Stolorz, P.E., Piatetsky-Shapiro, G. (eds.)
%  Proceedings of the Fourth International Conference on Knowledge Discovery and
%  Data Mining (KDD'98). pp. 274--278. AAAI Press, New York, NY, USA (1998)
\newblock KDD'98 (2008)

\bibitem[{Mizenmacher and Upfal(2005)}]{MitzenmacherU05}
  Mitzenmacher, M., Upfal, E.: Probability and Computing,  
  \newblock Cambridge University Press, second edn. (2005)

\bibitem[{Riondato and Upfal(2012)}]{RiondatoU12}
Riondato, M., Upfal, E.: Efficient discovery of association rules and frequent
  itemsets through sampling with tight performance guarantees.
%\newblock In: Flach, P.A., De~Bie, T., Cristianini, N. (eds.) Machine Learning
%  and Knowledge Discovery in Databases - European Conference, ECML PKDD 2012,
%  Bristol, UK, September 24-28, 2012. Proceedings, Part I. Lecture Notes in
% Computer Science, vol. 7523, pp. 25--41. Springer, Berlin / Heidelberg (2012)
\newblock \emph{To appear in} ACM Trans.~on Knowl.~Disc.~from Data  (2014)

\ifarxiv
\bibitem[{Riondato and Vandin(2014)}]{RiondatoV14}
Riondato, M., Vandin, F.: Finding the True Frequent Itemsets.
\newblock SDM'14 (2014)
\else
\bibitem[{Riondato and Vandin(2014)}]{RiondatoV14-full}
Riondato, M., Vandin, F.: Finding the True Frequent Itemsets.
\newblock CoRR abs/1301.1218v3 (2014).
\urlprefix\url{http://arxiv.org/pdf/1301.1218v3}
\fi

\bibitem[{Silverstein et~al.(1998)Silverstein, Brin, and
  Motwani}]{SilversteinBM98}
Silverstein, C., Brin, S., Motwani, R.: Beyond market baskets: Generalizing
  association rules to dependence rules.
\newblock Data Min.~and Knowl.~Disc.~2(1), 39--68 (1998)%,
  %\urlprefix\url{http://dx.doi.org/10.1023/A:1009713703947}

\iffalse
\bibitem[{Teytaud and Lallich(2001)}]{TeytaudL01}
Teytaud, O., Lallich, S.: Contribution of statistical learning to validation of
  association rules (2001)%, extended version of "Bornes unformes en extraction
 % de r\`{e}gles d'association", proceedings of CAp 2001
 \fi

\bibitem[{Vapnik and Chervonenkis(1971)}]{VapnikC71}
Vapnik, V.N., Chervonenkis, A.J.: On the uniform convergence of relative
  frequencies of events to their probabilities.
%\newblock Theory of Prob.~and its Appl.~16(2), 264--280 (1971)
\newblock Th.~Prob.~Appl.~16(2), 264--280 (1971)

\bibitem[{Webb(2007)}]{Webb07}
Webb, G.I.: Discovering significant patterns.
\newblock Mach.~Learn.~68(1), 1--33 (2007)

\bibitem[{Webb(2008)}]{Webb08}
Webb, G.I.: Layered critical values: a powerful direct-adjustment approach to
  discovering significant patterns.
\newblock Mach.~Learn.~71, 307--323 (2008)

\end{thebibliography}

