\section{The range set of a collection of itemsets}\label{sec:range}
In this section we define the concept of a range set associated to a
collection of itemsets and show how to bound the VC-dimension and the
empirical VC-dimension of this range set. We use these definitions and results
to develop our algorithm in later sections.

\begin{definition}\label{def:rangeset}
Given a collection $\mathcal{C}$ of itemsets built on a ground set $\Itm$, the
\emph{range set $\range(\mathcal{C})$ associated to $\mathcal{C}$ is a range
set on $2^\Itm$} containing the support sets of the itemsets in $\mathcal{C}$:
\[\range(\mathcal{C})=\{T(A) ~:~ A\in\mathcal{C}\}\enspace.\]
\end{definition}

\begin{fact}\label{fact:maxfreq}
	Given this definition and a dataset $\Ds$, the maximum
	from~\eqref{eq:evceapprox} is attained for the item $a\in\Itm$ with the
	highest frequency in $\Ds$, and the value of $|r\cap S|$ is exactly
	$f_\Ds(\{a\})|\Ds|$.
\end{fact}

The following Theorem presents an upper bound to the empirical VC-dimension of
$\range(\mathcal{C})$ on a dataset $\Ds$.

\begin{theorem}\label{lem:evcdimupbound}
  Let $\mathcal{C}$ be a collection of itemsets and let $\Ds$ be a dataset. Let
  $d$ be the maximum integer for which there are at least $d$
  transactions $\tau_1,\dotsc,\tau_d\in \Ds$ such that the set
  $\{\tau_1,\dotsc,\tau_d\}$ is an \emph{antichain}\footnote{An antichain is a
  collection of sets such no one of them is a subset of another.}, and each $\tau_i$, $1\le i\le d$,
  contains at least $2^{d-1}$ itemsets from $\mathcal{C}$.
  %\begin{enumerate*}
  %  \item
  %    The set $\{\tau_1,\dotsc,\tau_d\}$ is an antichain, and
  %    %For any pair of transactions $(\tau_i,\tau_j)$, $i\neq j$, we have
  %    %$\tau_i\not\subseteq\tau_j$ and $\tau_j\not\subseteq\tau_i$, and
  %  \item Each $\tau_i$, $1\le i\le d$ contains at least $2^{d-1}$ itemsets from
  %    $\mathcal{C}$.
  %\end{enumerate*}
  Then $\EVC(\range(\mathcal{C}),\Ds)\le d$.
\end{theorem}

\ifarxiv
\begin{proof}
  The antichain requirement guarantees that the set of transactions considered in
  the computation of $d$ could indeed theoretically be shattered. Assume that a
  subset $\mathcal{F}$ of $\Ds$ contains two transactions $\tau'$ and $\tau''$
  such that $\tau'\subseteq\tau''$. Any itemset from $\mathcal{C}$
  appearing in $\tau'$ would also appear in $\tau''$, so there would not be any
  itemset $A\in\mathcal{C}$ such that $\tau''\in T(A)\cap F$ but
  $\tau'\not\in T(A)\cap \mathcal{F}$, which would imply that $\mathcal{F}$ can
  not be shattered. Hence sets that are not antichains should not be
  considered. This has the net effect of potentially resulting in a lower $d$,
  i.e., in a stricter upper bound to $\EVC(\range(\mathcal{C}),\Ds)$.

  Let now $\ell>d$ and consider a set $\mathcal{L}$ of $\ell$ transactions from
  $\Ds$ that is an antichain. Assume that $\mathcal{L}$ is shattered by
  $\range(\mathcal{C})$. Let $\tau$ be a transaction in $\mathcal{L}$.
  The transactions $\tau$ belongs to $2^{\ell-1}$ subsets of $L$. Let
  $\mathcal{K}\subseteq \mathcal{L}$ be one of these subsets. Since
  $\mathcal{L}$ is shattered, there exists an itemset $A\in\mathcal{C}$ such
  that $T(A)\cap \mathcal{L}=\mathcal{K}$. From this and the fact
  that $t\in \mathcal{K}$, we have that $\tau\in T(A)$ or equivalently that
  $A\subseteq\tau$. Given that all the subsets $\mathcal{K}\subseteq\mathcal{L}$
  containing $\tau$ are different, then also all the $T(A)$'s such that
  $T(A)\cap \mathcal{L}=\mathcal{K}$ should be
  different, which in turn implies that all the itemsets
  $A$ should be different and that they should all appear in $\tau$. There are
  $2^{\ell-1}$ subsets $\mathcal{K}$ of $\mathcal{L}$ containing $\tau$,
  therefore $\tau$ must contain at least $2^{\ell-1}$ itemsets from
  $\mathcal{C}$, and this holds for all $\ell$ transactions in $\mathcal{L}$. This is a
  contradiction because $\ell>d$ and $d$ is the
  maximum integer for which there are at least $d$ transactions containing at
  least $2^{d-1}$ itemsets from $\mathcal{C}$. Hence $\mathcal{L}$ cannot be shattered and
  the thesis follows.
\end{proof}
\fi

\subsection{Computing the VC-Dimension}\label{sec:computvc}
The na\"ive computation of $d$  according to the definition in Thm.~\ref{lem:evcdimupbound}
requires to scan the transactions one by one,
compute the number of itemsets from $\mathcal{C}$ appearing in each
transaction, and make sure to consider only itemsets constituting antichains. Given the very large
number of transactions in typical dataset and the fact that the number of
itemsets in a transaction is exponential in its length, this method would be
computationally too expensive. An upper bound to $d$ (and therefore to
$\EVC(\range(\mathcal{C}),\Ds)$) can be computed by solving a
\emph{Set-Union Knapsack Problem} (SUKP)~\citep{GoldschmidtNY94} associated to
$\mathcal{C}$.
%where the elements to be put in the knapsack are organized into sets, which have a profit associated to them. The objective is to maximize
%the number of sets for which all the elements in those
%sets were put in the knapsack
%the profit, while satisfying the capacity constraint.

\begin{definition}[\citep{GoldschmidtNY94}]\label{def:sukp}
  Let $U=\{a_1,\dotsc,a_\ell\}$ be a set of elements and let
  $\mathcal{S}=\{A_1,\dotsc,A_k\}$ be a set of subsets of $U$, i.e.
  $A_i\subseteq U$ for $1\le i\le k$. Each subset $A_i$, $1\le i\le k$, has an associated
  non-negative \emph{profit} $\rho(A_i)\in\mathbb{R}^+$, and each element $a_j$, $1\le
  j\le\ell$ as an associated non-negative weight $w(a_j)\in\mathbb{R}^+$.
  Given a subset $\mathcal{S}'\subseteq\mathcal{S}$, we define the profit of
  $\mathcal{S}'$ as $P(\mathcal{S}')=\sum_{A_i\in \mathcal{S}'}\rho(A_i)$. Let
  $U_{\mathcal{S}'}=\cup_{A_i\in\mathcal{S}'} A_i$. We
  define the weight of $\mathcal{S}'$ as $W(\mathcal{S}')=\sum_{a_j\in
  U_{\mathcal{S}'}} w(a_j)$. Given a non-negative parameter $c$ that we call
  \emph{capacity}, the \emph{Set-Union Knapsack Problem} (SUKP) requires to find
  the set $\mathcal{S}^*\subseteq\mathcal{S}$ which \emph{maximizes}
  $P(\mathcal{S}')$ over all sets $\mathcal{S}'$ such that $W(\mathcal{S}')\le c$.
\end{definition}

In our case, $U$ is the set of items that appear in the itemsets of
$\mathcal{C}$, $\mathcal{S}=\mathcal{C}$, the profits and the weights are all
unitary, and the capacity constraint is an integer $\ell$. We call this
optimization problem the \emph{SUKP associated to $\mathcal{C}$ with capacity
$\ell$}.
It is easy to see %It should be evident
that the optimal profit of this SUKP is the maximum number
of itemsets from $\mathcal{C}$ that a transaction of length $\ell$ can contain.  %could fit into a transaction of length %$\ell$.
In order to show how to use this fact to compute an upper bound to
$\EVC(\range(\mathcal{C}),\Ds)$, we need to define some additional terminology. Let
$\ell_1,\dotsc,\ell_w$ be the sequence of the
\emph{transaction lengths} of $\Ds$, i.e., for each value $\ell$
for which there is at least a transaction in $\Ds$ of length $\ell$, there is
one (and only one) index $i$, $1\le i\le w$ such that $\ell_i=\ell$. Assume that
the $\ell_i$'s are labelled in sorted decreasing order:
$\ell_1>\ell_2>\dotsb>\ell_w$. Let now $L_i$, $1\le i\le w$ be the maximum number of
transactions in $\Ds$ that have length at least $\ell_i$ and such that
for no two $\tau'$, $\tau''$ of them we have either $\tau'\subseteq\tau''$ or
$\tau''\subseteq\tau'$. Let now $q_i$ be the optimal profit of the SUKP associated to
$\mathcal{C}$ with capacity $L_i$, and let $b_i=\lfloor \log_2q_i\rfloor +1$.
The sequences $(\ell_i)_1^w$ and a sequence $(L_i^*)^w$ of upper bounds to
$(L_i)_1^w$ can be computed efficiently with a scan of the dataset.
The following lemma uses these sequences to show how to obtain an upper bound to
the empirical VC-dimension of $\mathcal{C}$ on $\Ds$.

\begin{lemma}\label{lem:sukpevc}
  Let $j$ be the minimum integer for which $b_i\le L_i$. Then
  $\EVC(\mathcal{C},\Ds)\le b_j$. %We call $b_j$ the \emph{empirical b-index of
  %$\mathcal{C}$ on $\Ds$} and denote it as $\eb(\mathcal{C},\Ds)$
\end{lemma}
\ifarxiv
\begin{proof}
  If $b_j\le L_j$, then there are at least $b_j$ transactions which can contain
  $2^{b_j-1}$ itemsets from $\mathcal{C}$ and this is the maximum $b_i$ for
  which it happens, because the sequence $b_1,b_2,\dotsc,b_w$ is sorted in
  decreasing order, given that the sequence $q_1,q_2,\dotsc,q_w$ is. Then $b_j$
  satisfies the conditions of Lemma~\ref{lem:evcdimupbound}. Hence
  $\EVC(\mathcal{C},\Ds)\le b_j$.
\end{proof}
\fi

\begin{corollary}\label{lem:sukpvc}
  Let $q$ be profit of the SUKP associated to $\mathcal{C}$ with capacity
  equal to $\ell=|\{a\in\Itm ~:~ \exists A\in\mathcal{C} \mbox{ s.t. } a\in
  A\}|$ ($\ell$ is the number of items such that there is at least one itemset in $\mathcal{C}$ containing
  them).
  %$|\Itm|-1$.
  Let $b=\lfloor\log_2 q\rfloor + 1$. Then
  $\VC(\range(\mathcal{C}))\le b$. %We call $b*$ the \emph{b-index of
  %$\mathcal{C}$} and denote it as $\b(\mathcal{C})$.
\end{corollary}

\paragraph{Complexity and runtime considerations.} Solving the SUKP optimally is
NP-hard in the general case, although there are known restrictions for
which it can be solved in polynomial time using dynamic
programming~\citep{GoldschmidtNY94}. Since we have unit weights and unit
profits, our SUKP is equivalent to the \emph{densest $k$-subhypergraph} problem,
which can not be approximated within a factor of $2^{O(\log n)^\delta}$ for any
$\delta>0$ unless $3STA \in
DTIME(2^{n^{3/4+\varepsilon}})$~\citep{HajiaghayiJKLMRSV06}. A greedy algorithm
by~\citet{Arulselvan14} allows a constant factor approximation if each
items only appear in a constant fraction of itemsets of $\mathcal{C}$.
%It is not clear whether our case falls into
%one of these restrictions but we found that
For our case, it is actually \emph{not necessary to compute the optimal
solution} to the SUKP: any upper bound solution for which we can prove that
there is no power of two between that solution and the optimal solution would
result in the \emph{same upper bound} to the (empirical) VC-dimension, while
substantially speeding up the computation. This property can be specified in
currently available optimization problem solvers (e.g., CPLEX), which can then can compute the
bound to the (empirical) VC-dimension very fast even for very large instances
with thousands of items and hundred of thousands of itemsets in $\mathcal{C}$,
making this approach practical.

\paragraph{Refinements.} It is possible to make some refinements to our
computation of the \emph{empirical} VC-dimension of a collection $\mathcal{C}$ of
itemsets on a dataset $\Ds$. First of all, one can remove from $\mathcal{C}$ all
itemsets that never appear in $\Ds$, as the corresponding ranges can not help
shattering any set of transactions in $\Ds$. Identifying which itemsets to
remove requires a single linear scan of $\Ds$. Secondly, when computing the
capacities $L_i$ (i.e., their upper bounds $L_i^*$), we can ignore all the
transactions that do not contain \emph{any} of the itemsets in $\mathcal{C}$ (or
the filtered version of $\mathcal{C}$), as there is no way of shatter them using
the ranges corresponding to itemsets in $\mathcal{C}$. Both refinements aim at
reducing the optimal value of the SUKP associated to $\mathcal{C}$, and
therefore at computing a smaller bound to the empirical VC-dimension of
$\mathcal{C}$ on $\Ds$. We remark that these refinements can not be used when
computing the (non-empirical) VC-dimension.

\paragraph{The range set of all itemsets.}
The range set associated to $2^\Itm$
is particularly interesting for us. It is possible  %It is easy
to compute bounds to $\VC(\range(2^\Itm))$ and
$\EVC(\range(2^\Itm), \Ds)$ without having to solve a SUKP.
%\citet{RiondatoU12} gave an upper bound to the empirical VC-dimension of
%$\range(2^\Itm)$ on $\Ds$.
%Before presenting the bound, we need to introduce a
%characteristic quantity of the set of transactions $\Ds$ called the \emph{d-index} and
%denoted as $\mathsf{d}(\Ds)$.

%\begin{definition}\label{def:dindex}
\begin{theorem}[\citep{RiondatoU12}]\label{thm:empvcdimubfirst}
  Let $\Ds$ be a dataset built on a
  ground set $\Itm$. The
  \emph{d-index} $\mathsf{d}(\Ds)$ of $\Ds$ is the maximum integer $d$ such that
  $\Ds$ contains at least $d$ transactions of length at least $d$ %such that
  %no one of them is a subset of another, that is, they form an antichain
  that form an antichain. %
  %for any two of these transactions $\tau',\tau''$ we have neither $\tau'\subseteq
  %\tau''$ nor $\tau''\subseteq \tau'$.
  We have $\EVC(\range(2^\Itm),\Ds)\le \mathsf{d}(\Ds)$.
%\end{definition}
\end{theorem}

%\citet{RiondatoU12} presents an efficient algorithm to compute an upper bound to
%the d-index of a dataset with a single linear scan of the dataset $\Ds$.

%\begin{theorem}\label{thm:empvcdimubfirst}
%  \emph{(\citep[Thm.~3]{RiondatoU12})} Let $\Ds$ be a dataset  with
%  transactions built on a ground set $\Itm$. Then $\EVC(\range(2^\Itm),\Ds)\le
%  \mathsf{d}(\Ds)$.
%\end{theorem}

%It easy to extend this theorem to prove an upper bound to the VC-dimension of
%$\range(2^\Itm)$.
\begin{corollary}\label{thm:vcdimubfirst}
  %Let $\Ds$ be a dataset with transactions built on a ground set $\Itm$. Then
  $\VC(\range(2^\Itm))\le |\Itm|-1$.
\end{corollary}
%\begin{proof}
%  Let $k=|\Itm|$. Assume w.l.o.g.~that $\Itm=\{1,\dotsc,k\}$. Let
%  $\tau_i=\Itm\setminus\{i\}$, $1\le i\le k$. We have $|\tau_i|=k-1$, for $1\le
%  i\le k$. Notice that for $i\neq j$, neither $\tau_i\subseteq\tau_j$ nor
%  $\tau_j\subseteq\tau_i$. Consider a dataset $\Ds$ made of the set of
%  transactions $\{\tau_1,\dotsc,\tau_k\}$ and any number of additional
%  transactions. It is easy to see that $\mathsf{d}(\Ds)=k-1$, and that
%  it is impossible to build a dataset with a larger d-index using the items of
%  $\Itm$, because there is no way to use the items of $\Itm$ to build $\ell>k-1$
%  transactions of length at least $\ell$ and such that for any two of these
%  transactions $\tau',\tau''$ we have neither $\tau'\subseteq \tau''$ nor
%  $\tau''\subseteq \tau'$. This concludes the proof. \qed
%\end{proof}

\citet{RiondatoU12} presented an efficient algorithm to compute an upper bound to
the d-index of a dataset with a single linear scan of the dataset $\Ds$.
The upper bound presented in Thm.~\ref{thm:empvcdimubfirst} is tight:
%, in the sense that
 there are datasets for which
$\EVC(\range(2^\Itm),\Ds)=\mathsf{d}(\Ds)$~\citep{RiondatoU12}.
This implies that the upper bound presented in Corol.~\ref{thm:vcdimubfirst} is
also tight.

%The upper bound to the empirical VC-Dimension of $\range(2^\Itm)$ on $\Ds$ is used
%by~\citet{RiondatoU12} to compute the size of a sample $\Sam$ needed to extract
%a superset of the collection of $\FI(\Ds,\Itm,\theta)$ from $\Sam$, where the
%quality of the approximation is controlled by user-defined parameters
%$\varepsilon$ and $\delta$. In this work we will use the two upper bounds
%presented above to compute a superset of the negative border of the collection
%of RFI's with respect to a minimum real frequency threshold $\theta$.

